New insights into the Weddell Sea ecosystem applying a network approach

### 2022-08-10 Manuscript/Manuscript.tex

Finally, we run extinction simulations and estimated its impact on the
stability of the network. For this, we calculated a stability index
called Quasi-Sign Stability (QSS), which is the proportion of stable
networks using randomized Jacobians and keeping the predator-prey sign
structure fixed \citep{Allesina2008}. With the aim of analysing the
effect of each species on the food web's stability, we deleted one
species at a time, so the network size was reduced by one. After each
species extinction, we calculated the QSS for the food web minus one
species (size = 489) and compared it with the QSS for the whole network
(size = 490); we performed 1000 simulations for each species extinction.
Then we statistically analysed such difference in QSS with an
Anderson-Darling test considering a p-value \textless{} 0.01
\citep{Scholz1987}. The formula for the QSS is described in
Supplementary Material.

@@

Figure 6

Biplots btw QSS difference and weighted and unweighted prop.

@@

\begin{table}[t]
\caption{Model comparison for the distribution of interaction strengths of the Weddell Sea food web. According to AIC, log-Normal is the best model.}
\begin{tabular}{l c c c}
\tophline

Model & df & AIC & deltaAIC \\
\middlehline
log-Normal & 2 & -356579 & 0 \\
\middlehline
Gamma & 2 & -352575 & 4004 \\
\middlehline
Power-law & 2 & -347646 & 8933 \\
\middlehline
Exponential & 1 & -262852 & 93726 \\
\middlehline
Normal & 2 & -222785 & 133793 \\
\middlehline
Uniform & 2 & -178001 & 178578 \\

>>>>

Finally, we run extinction simulations and estimated its impact on the
stability of the network. For this, we calculated a stability index
called Quasi-Sign Stability (QSS), which is the proportion of stable
networks using randomized Jacobians and keeping the predator-prey sign
structure fixed \citep{Allesina2008}. If this proportion is zero, then
one should take into account the real part of the maximum eigenvalue of
the Jacobian matrix, which is also a measure of stability
\citep{Grilli2016}.

With the aim of analysing the effect of each species on the food web's
stability we deleted one species at a time, so the network size was
reduced by one. After each species extinction, we calculated the
stability for the food web minus one species (size = 489) and compared
it with that of the whole network (size = 490). We performed 1000
simulations for each species extinction and obtained a mean QSS and
maximum eigenvalue. At last we statistically analysed such difference
with an Anderson-Darling test considering a p-value \textless{} 0.01
\citep{Scholz1987}. If the difference was positive, then the stability
of the food web was impacted positively (higher stability) when that
species become extinct, and viceversa. Details for the stability
calculations are described in Supplementary Material.

Once we had the results for the impact on stability for each species
extinction, we plotted them considering weighted (interaction strength)
and unweighted properties, and species habitat.

@@

It's important to note that since the proportion of Jacobians that are
locally stable or QSS was zero for the Weddell Sea food web, we
considered the mean maximum eigenvalue as the stability index. Our
results showed that the majority of the species had no significant
impact on the stability of the food web when extinct (Anderson-Darling
test p-value \textgreater{} 0.01). This is clearly observed in Figure 6
A-D, where most of the points lie around the zero value of the
`Stability difference' axis. This means that the stability for the food
web minus a given species is similar to the stability for the whole
network. However, there were few species that when become extinct
generated positive significant differences in the stability of the food
web, meaning that they increased the network's stability when deleted.
There is an exception, Pagetopsis macropterus (demersal fish), that when
deleted decreased the stability. In total, 13 out of 490 species
(2.65\%) gave rise to significant changes in the network's stability.

When plotting the stability difference against species properties and
the previous results regarding clustering, interesting insights arose
(Figure 6). In the first place, all species with a significant impact on
the stability belong to the `High IS' group, denoting a high mean
interaction strength (Figure 6A). Secondly, these species are positioned
in high trophic levels (TL \textgreater{} 3.7) (Figure 6B). Thirdly,
they have a relatively high number of interactions or degree (Degree
\textgreater{} 25) (Figure 6C). Regarding trophic similarity, species
have mid to low values (TS \textless{} 0.13) (Figure 6D). Habitat wise,
species live in all the habitats, except for are benthic (Figure 6E).
Table 2 shows these results per species.

@@

\begin{table}[t]
\caption{Model comparison for the distribution of interaction strengths of the Weddell Sea food web. Order by best fit. References: df = degrees of freedom, AIC = Akaike Information Criterion, deltaAIC = difference with best fit. Log-Normal is the best model.}
\begin{tabular}{l c c c}
\tophline

\textbf{Model} & \textbf{df} & \textbf{AIC} & \textbf{deltaAIC} \\
\middlehline
log-Normal & 2 & -356579 & 0 \\
\middlehline
Gamma & 2 & -352575 & 4004 \\
\middlehline
Power-law & 2 & -347646 & 8933 \\
\middlehline
Exponential & 1 & -262852 & 93726 \\
\middlehline
Normal & 2 & -222785 & 133793 \\
\middlehline
Uniform & 2 & -178001 & 178578 \\

\begin{table}[t]
\caption{Properties of the species that when become extinct generated a significant impact on the stability of the Weddell Sea food web, ordered by significance (Anderson-Darling p value). All species belong to the 'High IS' group.  References: meanIS = mean interaction strength, TL = trophic level, Deg = degree, TS = trophic similarity, StabDif = stability difference, ADvalue = Anderson-Darling p value.}
\begin{tabular}{l c c c c c c c}
\tophline

\textbf{Species} & \textbf{meanIS} & \textbf{TL} & \textbf{Deg} & \textbf{TS} & \textbf{Habitat} & \textbf{StabDif} & \textbf{ADvalue}\\
\middlehline
Orcinus orca & 0.00163 & 5.03 & 26 & 0.037 & Benthopelagic & 0.00040 & 1.38e-51 \\
\middlehline
Macrourus holotrachys & 0.00074 & 4.70 & 85 & 0.112 & Benthopelagic & 0.00027 & 2.75e-18 \\
\middlehline
Abyssorchomene nodimanus & 0.00022 & 4.21 & 137 & 0.130 & Benthopelagic & 0.00024 & 1.02e-15 \\
\middlehline
Dissostichus mawsoni & 0.00069 & 4.12 & 87 & 0.126 & Pelagic & 0.00018 & 1.06e-10 \\
\middlehline
Pagetopsis macropterus & 0.00063 & 4.64 & 76 & 0.113 & Demersal & -0.00017 & 1.66e- 9 \\
\middlehline
Macrourus whitsoni & 0.00064 & 4.55 & 92 & 0.124 & Benthopelagic & 0.00012 & 4.97e- 8 \\
\middlehline
Hydrurga leptonyx & 0.00092 & 4.72 & 67 & 0.094 & Land-based & 0.00014 & 5.59e- 6 \\
\middlehline
Mesonychoteuthis hamiltoni & 0.00161 & 4.41 & 29 & 0.028 & Pelagic & 0.00013 & 6.53e- 6 \\
\middlehline
Notothenia marmorata & 0.00073 & 4.09 & 44 & 0.091 & Demersal & 0.00014 & 2.10e- 5 \\
\middlehline
Champsocephalus gunnari & 0.00068 & 3.72 & 46 & 0.086 & Pelagic & 0.00010 & 7.97e- 5 \\
\middlehline
Arctocephalus gazella & 0.00083 & 4.67 & 61 & 0.093 & Land-based & 0.00011 & 9.52e- 5 \\
\middlehline
Ommatophoca rossii & 0.00093 & 4.87 & 56 & 0.080 & Land-based & 0.00009 & 1.20e- 3 \\
\middlehline
Notothenia coriiceps & 0.00044 & 4.27 & 130 & 0.126 & Demersal & 0.00012 & 1.33e- 3 \\
